\documentclass[11pt]{article}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{hyperref} 
\usepackage{enumerate}

\oddsidemargin 0mm
\evensidemargin 5mm
\topmargin -20mm
\textheight 240mm
\textwidth 160mm



\pagestyle{myheadings} 
\markboth{Homework 1}{Spring 2016 CS 4269/6362 Machine Learning: Homework 1} 


\title{CS 269/362 Machine Learning: Homework 1\\Learning Foundations\\
\Large{Due: 1/19/2016}\\
50 Points Total \hspace{1cm} Version 1.0}
\author{}
\date{}

\begin{document}
\large
\maketitle
\thispagestyle{headings}

\vspace{-.5in}

\paragraph{1 (3 points)} Explain why you agree or disagree with the following statement.\\
{\it It is always best to select a hypothesis class that contains the optimal hypothesis.}

\paragraph{2 (3 points)} True or False: {\it An infinite hypothesis class always contains the optimal hypothesis.} If true, why? If false, give a counter example.

\paragraph{3 (3 points)} Consider the following development scenario. A researcher collects 1000 labeled examples for learning, dividing them into a training set (500 examples) and a test set (500 examples). To develop a classifier, the researcher experiments by adding new features. To guide feature construction, the research tests each set of features by training on the train data and testing on the test data, measuring the resulting change in accuracy and only keeping features that help. When the researcher is finished, he collects 1000 new labeled examples and evaluates the classifier on these new examples.

Do you expect accuracy on these 1000 new examples to be the same, better or worse than the original 500 test examples? Why?

\paragraph{4 (3 points)} At the start of the semester, you arrive home to find three packages containing your three new textbooks. Each package was sent via Fedex or UPS with equal probability. Of the three packages, one of the packages is a brown box delivery by UPS. What is 
the probability that you have one UPS package and two Fedex packages? Why?

\paragraph{5 (3 points)} 
True/False (and why): Suppose you know your hypothesis class contains the optimal hypothesis, and you observe that changing any one small part (e.g., a single weight) of your current hypothesis makes it worse than before. You can safely conclude that the current hypothesis must be optimal.

\end{document}